### 2017-02-21 2017mergingmerges/section/XX-Processes.tex
ok
fig 5. fusion of group and merge

Fig.~\ref{fig:Process:Fused} shows the result of fusing @group@ and @merge@ together.
The process has two inputs, @s1@ and @s2@, which correspond to the original input streams.
Of the two outputs, @s3@ is the @group@'s output stream, while @s4@ is the @merge@'s output stream.
The heap contains the variables from both processes, as well as the new buffer variable @b1@, which has been initialised to $0$.

\begin{figure}
\begin{code}
            process
            { ins:    { s1, s2 }
            , outs:   { s3, s4 }
            , heap:   { f = T, l = 0, v = 0, x1 = 0, x2 = 0, b1 = 0}
            , label:  F0
            , instrs:

  (shared pull)
              { F0  = pull s1 b1            F1  {}

  (merge label B0)
              , F1  = jump                  F2  {v  = b1}

  (group labels A0-A3)
              , F2  = jump                  F3  {x1 = b1}
              , F3  = case (f || (l /= v))  F4  {}                F5  {}
              , F4  = push s3 v             F5  {l  = v, f  = F}
              , F5  = jump                  F6  {}

  (merge label B1)
              , F6  = pull s2 x2            F7  {}

  (merge label C0)
              , F7  = case (x1 < x2)        F8  {}                F16 {}

  (merge labels D0-D1)
              , F8  = push s4 x1            F9  {}
              , F9  = drop s1               F10 {}

  (shared pull)
              , F10 = pull s1 b1            F11 {}

  (group labels A0-A3)
              , F11 = jump                  F12 {v  = b1}
              , F12 = case (f || (l /= v))  F13 {}                F14 {}
              , F13 = push s3 v             F14 {l  = v, f  = F}
              , F14 = jump                  F15 {}

  (merge label D2)
              , F15 = jump                  F7  {x1 = b1}

  (merge labels E0-E2)
              , F16 = push s4 x2            F17 {}
              , F17 = drop s2               F18 {}
              , F18 = pull s2 x2            F7  {} } }
\end{code}

>>>>

Fig.~\ref{fig:Process:Fused} shows the entire result of fusing @group@ and @merge@ together.
The process has two inputs, @s1@ and @s2@, which correspond to the original input streams.
Of the two outputs, @s3@ is the @group@'s output stream, while @s4@ is the @merge@'s output stream.
The heap contains the variables from both processes, as well as the new buffer variable @b1@, which is used to store the current element pulled from @s1@.

For the most part, the fused process simply alternates between performing the instructions of @group@ and those of @merge@.
For the shared input stream @s1@, the two processes must be coordinated so that values are only pulled or dropped when both agree.

The instructions of @merge@ can be split into four categories, which can be identified in the fused process: labels @B0@-@B2@ perform initialisation, and are mapped to @F0@-@F6@; label @C0@ performs case analysis to find the smaller value and is mapped to @F7@; labels @D0@-@D2@ push the value from the first stream, @s1@, and are mapped to @F8@-@F15@; and labels @E0@-@E2@ push the value from the second stream, @s2@, and are mapped to @F16@-@F18@.

The @group@ process can also be identified in the fused process: the fused labels @F0@-@F5@ perform the pulling, case analysis and pushing from instructions @A0@-@A2@.
These instructions are seen again in @F10@-@F14@, when the @merge@ process is pulling from the @s1@ stream.
Thus the @group@ instructions have been duplicated, and are performed once when @merge@ initialises, and again every time @merge@ pulls from the first stream.

Labels @F5@ and @F14@ correspond to the (@drop s1@) instruction at @A3@ in @group@, but note that they are fused as @jump@ instructions.
As @s1@ is shared between the two processes, the element from @s1@ can only be dropped once both processes agree to drop.
In the fused labels on the right hand column, the other process still has @s1@ as `pending' or `have', so the element cannot yet be dropped.
The actual @drop@ is only performed once both processes are `none', at label @F9@.

\begin{figure}
\begin{code}
process
{ ins:    { s1, s2 }
, outs:   { s3, s4 }
, heap:   { f = T, l = 0, v = 0, x1 = 0, x2 = 0, b1 = 0}
, label:  F0
, instrs:
\end{code}

\newcommand\annot[5]{
  \tiny ((#1, \> \tiny \{s1 =      #2\}), \> \tiny (#3, \> \tiny \{s1 =      #4, \> \tiny s2 =      #5\}))
}
\newcommand\icase[7]{
 \tt{#1} \> \tt{= #2} \> \tt{#3} \{\tt{#4}\} \> \tt{#5} \{\tt{#6}\} \> \tiny #1 \> \tiny = #7 \\
}
\newcommand\instr[5]{
 \tt{#1}\>\tt{= #2} \> \tt{#3} \{\tt{#4}\} \> \> \tiny #1 \> \tiny = #5 \\
}
\begin{tabbing}
@  @ \=
@, F17@ \= = @case (f || (l /= v)) @
      \= @F17{x1 = b1, f = F}@ \= @F17{ }@
@      @ \= \tiny F17 \= \tiny = ((A0, \= \tiny \{s1 = pending\}), \= \tiny (B0, \= \tiny \{s1 = pending, \= \tiny s2 = pending\})) \kill
\> @{@ \instr{F0}{pull s1 b1}{F1}{}{\annot{A0}{none}{B0}{none}{none}}

\> @,@
\instr{F1}{jump}{F2}{v  = b1}
      {\annot{A0}{pending}{B0}{pending}{none}}
\> @,@
\instr{F2}{jump}{F3}{x1 = b1}
      {\annot{A1}{have}{B0}{pending}{none}}
\> @,@
\icase{F3}{case (f || (l /= v))}{F4}{}{F5}{}
      {\annot{A1}{have}{B1}{have}{none}}

\> @,@
\instr{F4}{push s3 v}{F5}{l = v, f = F}
      {\annot{A2}{have}{B1}{have}{none}}

\> @,@
\instr{F5}{jump}{F6}{}
      {\annot{A3}{have}{B1}{have}{none}}

\> @,@
\instr{F6}{pull s2 x2}{F7}{}
      {\annot{A0}{none}{B1}{have}{none}}
\\

\> @,@
\icase{F7}{case (x1 < x2)}{F8}{}{F16}{}
      {\annot{A0}{none}{C0}{have}{have}}

\\
\> @,@
\instr{F8}{push s4 x1}{F9}{}
      {\annot{A0}{none}{D0}{have}{have}}
\> @,@
\instr{F9}{drop s1}{F10}{}
      {\annot{A0}{none}{D1}{none}{have}}
\> @,@
\instr{F10}{pull s1 b1}{F11}{}
      {\annot{A0}{none}{D2}{none}{have}}
\> @,@
\instr{F11}{jump}{F11}{v = b1}
      {\annot{A0}{pending}{D2}{pending}{have}}
\> @,@
\icase{F12}{case (f || (l /= v))}{F13}{}{F14}{}
      {\annot{A1}{have}{D2}{pending}{have}}
\> @,@
\instr{F13}{push s3 v}{F14}{l = v, f = F}
      {\annot{A2}{have}{D2}{pending}{have}}
\> @,@
\instr{F14}{jump}{F15}{}
      {\annot{A3}{have}{D2}{pending}{have}}
\> @,@
\instr{F15}{jump}{F7}{x1 = b1}
      {\annot{A0}{none}{D2}{pending}{have}}

\\
\> @,@
\instr{F16}{push s4 x2}{F17}{}
      {\annot{A0}{none}{E0}{have}{have}}
\> @,@
\instr{F17}{drop s2}{F18}{}
      {\annot{A0}{none}{E1}{have}{have}}
\> @,@
\instr{F18}{pull s2}{F7}{}
      {\annot{A0}{none}{E2}{have}{none}}

@} }@
\end{tabbing}
